---
title: "STA206_ASS3"
author: "Zhen Zhang"
date: "October 14, 2015"
output: pdf_document
---

```{r global_options, include=FALSE, echo=FALSE, cache=FALSE}
options(width=70)
knitr::opts_chunk$set(warning=FALSE, message=FALSE, tidy=TRUE, fig.path='Figs/')
library(Matrix)
```

(a) Model Equations: $y = \beta_{0} + \beta_{1} * X_1 + \beta_{2} * X_2 + \epsilon$

Coefficient Vector: $\beta = [\beta_0, \beta_1, \beta_2]$

Design Matrix: $X = [\mathbf{1}, X_1, X_2]$

Response Vector: $y$

(b) First read data into R

```{r}
X1 <- c(-0.63, 0.18, -0.84, 1.60, 0.33)
X2 <- c(-0.82, 0.49, 0.74, 0.58, -0.31)
Y  <- c(-0.97, 2.51, -0.19, 6.53, 1.00)
```

Then create design matrix $\mathbf{X}$ and the response vector $\mathbf{Y}$

```{r}
(X <- matrix(c(rep(1,5), X1, X2), 5, 3))
```

Now calculate $\mathbf{X'X}$, $\mathbf{X'Y}$ and $\mathbf{(X'X)^{-1}}$

```{r}
# XX value
(XX <- t(X) %*% X)
# XY value
(XY <- t(X) %*% Y)
# XX Inverse
(XXInv <- solve(XX))
```

(c) The least square estimator $\hat{\beta}$ is

```{r}
(beta <- XXInv %*% XY)
```

(d) The hat matrix $\mathbf{H}$ is

```{r}
(H <- X %*% XXInv %*% t(X))
```

And the rank is

```{r}
rankMatrix(H)
rankMatrix(diag(5) - H)
```

(e)

```{r}
(Y_fit <- X %*% beta)
(e <- Y - Y_fit)
(SSE <- t(e) %*% e)
```

Since SSE has degree of $5-2-1 = 2$, so

```{r}
(MSE <- SSE / 2)
```

(f) Model Equations: $y = \beta_{0} + \beta_{1} * X_1 + \beta_{2} * X_2 + \beta_3 * X_1X_2 + \epsilon$

Coefficient Vector: $\beta = [\beta_0, \beta_1, \beta_2, \beta_3]$

(g) First create the new design matrix $X_{Int}$

```{r}
(X_Int <- matrix(c(rep(1,5), X1, X2, X1 * X2), 5, 4))
```

The response vector, $Y_{Int}$, is unchanged:

```{r}
(Y_Int <- Y)
```

Hat Matrix $H_{Int}$

```{r}
(H_Int <- X_Int %*% solve(t(X_Int) %*% X_Int) %*% t(X_Int))
```

Now calculate the rank:

```{r}
rankMatrix(H_Int)
rankMatrix(diag(5) - H_Int)
```

The rank of $H$ is now 4, which is the number of $\beta$

(h) The $\beta_{Int}$ here is

```{r}
beta_Int <- solve((t(X_Int) %*% X_Int)) %*% (t(X_Int) %*% Y_Int)
```

(i) Fitted values:

```{r}
(Y_fit_Int <- X_Int %*% beta_Int)
(e_Int <- Y_Int - Y_fit_Int)
(SSE_Int <- t(e_Int) %*% e_Int)
```

The degree of freedom for SSE here is $5-3-1 = 1$

```{r}
(MSE_Int <- SSE_Int / 1)
```

(j) The second model (the model with interaction) fits the data better, since its MSE and SSE is smaller than the previous one.