---
title: "STA207 homework4"
author: "Juanjuan Hu, Zhen Zhang"
date: "February 7, 2016"
output: pdf_document
---
# 25.7
## (a)
```{r}
library(xlsx)
setwd("/Users/PullingCarrot/Desktop/201601-03/sta207/homework/hw4")
sodium = read.xlsx2("DataFor+PR25.7.xlsx",1)
sodium$Sodium = as.numeric(as.character(sodium$Sodium))
sodium$Brand = as.factor(sodium$Brand)
result = aov(Sodium ~ Brand, data=sodium)
summary(result)
```
$H_0: {\sigma_\mu}^2 = 0$

$H_1: {\sigma_\mu}^2 > 0$

$F^*=\frac{MSTR}{MSE}=170.91/0.72=237.375$

$F(0.99, 5, 42) = 3.49$

$p-val = 0$

The decision rule is: if $F^*$ is greater than 3.49, then reject $H_0$, otherwise, accept $H_0$.
Here, $237.375 > 3.49$, so we reject $H_0$, concluding that the mean sodium content is not the same in all brands sold in the metropolitan area. The p-value is almost zero, which is less than 0.01, leading to the same conclusion.

## (b)
```{r}
sum(sodium$Sodium)/nrow(sodium)
```
$\hat{\mu}_{.} = \bar{Y}_{..}=17.63$

$s^2 = MSTR/r*n = 170.91/48 = 3.560625$

$s(\bar{Y}_{..}) = \sqrt{3.560625}=1.887$

$t(0.995, 5)=4.032$

The confidence interval is $\bar{Y}_{..}\pm t(1-\alpha, r-1)*s(\bar{Y}_{..})$, which is $17.63 \pm 4.032*1.887$. i.e. [10.021, 25.237].

# 25.8
## (a)
According to (25.16), compute L and U.

$F(\alpha/2; r-1, r(n-1)) = F(0.005, 5, 42) = 0.0799$

$F(1-\alpha/2; r-1, r(n-1)) = F(0.995, 5, 42) = 3.9528$

$L = \frac{1}{n}(\frac{MSTR}{MSE}*\frac{1}{F(1-\alpha/2; r-1, r(n-1))}-1) = \frac{1}{8}(\frac{170.91}{0.72}*\frac{1}{F(0.995, 5, 42)}-1)=7.4292$

$U = \frac{1}{n}(\frac{MSTR}{MSE}*\frac{1}{F(\alpha/2; r-1, r(n-1))}-1) = \frac{1}{8}(\frac{170.91}{0.72}*\frac{1}{F(0.005, 5, 42)}-1)=375.2083$

$L^* = \frac{L}{1+L} = 7.4292/(1+7.4292) = 0.8814$

$U^* = \frac{U}{1+U} = 375.2083/(1+375.2083) = 0.9973$

Therefore, the 99% confidence interval for $\frac{(\sigma_{\mu})^2}{(\sigma_{\mu})^2+(\sigma)^2}$ is [0.8814, 0.9973]. It means that with confidence coefficient 0.99, we conclude that the variability of the mean sodium for the different brands accounts for somewhere between 88 and 99.8 percent of the total variability of the sodium levels.

## (b)
According to (25.6), an unbiased estimator of $\sigma^2$ is MSE = 0.72. According to (25.23), an unbiased point estimator of ${\sigma_{\mu}}^2$ is ${s_{\mu}}^2 = \frac{MSTR - MSE}{n} = (170.91-0.72)/8=21.2738$.

## (c)
$\chi^2(1-\alpha/2;42) = \chi^2(0.995;42)=69.336$

$\chi^2(\alpha/2;42) = \chi^2(0.005;42)=22.1385$

According to (25.21), the CI is: $\frac{r(n-1)MSE}{\chi^2(1-\alpha/2;42)}\leq \sigma^2 \leq \frac{r(n-1)MSE}{\chi^2(\alpha/2;42)}$. Here the 99% confidence interval is: $\frac{6(8-1)0.72}{\chi^2(0.995;42)}\leq \sigma^2 \leq \frac{6(8-1)0.72}{\chi^2(0.005;42)}$, i.e. [0.4361, 1.3659].

## (d)
$H_0: {\sigma_{\mu}}^2 \leq 2\sigma^2$

$H_1: {\sigma_{\mu}}^2 > 2\sigma^2$

$F^*=\frac{MSTR/2n+1}{MSE}=(170.91/17)/0.72=13.96$

$F(0.99, 5, 42) = 3.49$

The decision rule is: if $F^*$ is greater than 3.49, then reject $H_0$, otherwise, accept $H_0$.
Here, $13.96 > 3.49$, so we reject $H_0$, concluding that the variance of sodium content between brands is more than twice as great as that within brands.

## (e)
$c_1 = 1/n = 1/8=0.125$, $c_2 = -c_1 = -0.125$

$\hat{L} = c_1MSTR +c_2MSE = 0.125*170.91-0.125*0.72=21.2737$

$df_1 = r-1 =5$, $df_2 = r(n-1) = 6*7 = 42$

$F_1 = F(1-\alpha/2, df_1, \inf) = F(0.995,5,\inf) =3.35$

$F_2 = F(1-\alpha/2, df_2, \inf) = F(0.995,42,\inf) =1.66$

$F_3 = F(1-\alpha/2, \inf, df_1) = F(0.995,\inf,5) =12.1$

$F_4 = F(1-\alpha/2, \inf, df_2) = F(0.995,\inf,42) =1.91$

$F_5 = F(1-\alpha/2, df_1, df_2) = F(0.995,5,42) =3.95$

$F_1 = F(1-\alpha/2, df_2, df_1) = F(0.995,42,5) =12.51$

$G_1 = 1-\frac{1}{F_1} = 1-1/3.35=0.7015$

$G_2 = 1-\frac{1}{F_2} = 1-1/1.66=0.3976$

$G_3 = \frac{(F_5-1)^2-(G_1F_5)^2-(F_4-1)^2}{F_5} = 0.0497$

$G_4 = F_6((\frac{F_6-1}{F_6})^2-(\frac{F_3-1}{F_6})^2 -{G_2}^2)=-1.2371$

$H_L = 14.99$

$H_U = 237.127$

According to (25.34), the 99% confidence interval is $\hat{L} - H_L \leq L \leq \hat{L} + H_U$, which is [21.2737-14.99, 21.2737+237.127], i.e. [6.284, 258.401]. It means that with confidence coefficient 0.99, we conclude that the variability of the mean sodium between brands is somewhere between 6.28 and 258.4.

# 25.16
## (a)
```{r}
setwd("/Users/PullingCarrot/Desktop/201601-03/sta207/homework/hw4")
disk = read.xlsx2("DataForPR25.16.xlsx",1)
sapply(disk, class)
disk$Minutes = as.numeric(as.character(disk$Minutes))
model = aov(Minutes ~ Make + Error(Make*Technician), data = disk)
summary(model)
```
$H_0: {\sigma_{\alpha\beta}}^2 = 0$

$H_1: {\sigma_{\alpha\beta}}^2 > 0$

$F^*=\frac{MSAB}{MSE}=303.8/52.01=5.8412$

$F(0.99, 4, 36) = 3.89$

$p-val = 0.0009944268$

The decision rule is: if $F^*$ is greater than 3.89, then reject $H_0$, otherwise, accept $H_0$.
Here, $5.8412 > 3.49$, so we reject $H_0$, concluding that the two factors interact. The p-value is nearly 0.001, which is less than 0.01, leading to the same conclusion.

## (b)
$MSAB=303.8$, $MSE = 52.01$

The point estimator of ${\sigma_{\alpha \beta}}^2$ is $\frac{MSAB-MSE}{n}=\frac{303.8-52.01}{5}=50.358$, which is no larger than $\sigma^2$.

## (c)
$H_0: {\sigma_{\alpha}}^2 = 0$

$H_1: {\sigma_{\alpha}}^2 > 0$

$F^*=\frac{MSA}{MSE}=12.29/52.01=0.2363$

$F(0.99, 2, 36) = 5.25$

$p-val = 0.7907587$

The decision rule is: if $F^*$ is greater than 5.25, then reject $H_0$, otherwise, conclude $H_0$.
Here, $0.2363 < 5.25$, so we cannot reject $H_0$, concluding that the factor A main effects are not present. The p-value is 0.79, which is above 0.01, leading to the same conclusion. It is meaningful to test this becuase if we can reject $H_0$, saying no A effects, then we can simplify our model.

## (d)
$H_0$: all $\beta_j = 0$

$H_1$: not all $\beta_j$ equals 0

$F^*=\frac{MSB}{MSAB}=14.16/303.8=0.0466$

$F(0.99, 2, 4) = 18$

$p-val = 0.9549795$

The decision rule is: if $F^*$ is greater than 18, then reject $H_0$, otherwise, conclude $H_0$.
Here, $0.0466 < 18$, so we cannot reject $H_0$, concluding that the factor B main effects are not present. The p-value is 0.955, which is above 0.01, leading to the same conclusion.It is meaningful to test this becuase if we can reject $H_0$, saying no B effects, then we can simplify our model.

## (e)
```{r}
mean(subset(disk, Make==1)$Minutes)
mean(subset(disk, Make==2)$Minutes)
mean(subset(disk, Make==3)$Minutes)
```
$\bar Y_{.1.} = 56.13$, $\bar Y_{.2.} = 56.6$,$\bar Y_{.3.} = 54.73$

$\hat D_1=\bar Y_{.1.}-\bar Y_{.2.}=-0.47$, $\hat D_2=\bar Y_{.1.}-\bar Y_{.3.}=1.4$,$\hat D_3=\bar Y_{.2.}-\bar Y_{.3.}=1.87$

$s(\hat{D_i})=\sqrt{\frac{MSAB}{bn}*2}=\sqrt{\frac{303.8}{5*3}*2}=6.3645$

$q(0.95,3,4) = 5.04$

$T = q(0.95,3,4)*(1/\sqrt{2})=3.5638$

Therefore, using the Turkey procedure and a 0.95 family confidence coefficient, the confidence intervals are:

$\hat D_1-s(\hat {D_1})*T \leq D_1 \leq \hat D_1+s(\hat{D_1})*T$, i.e. $-23.15 \leq D_1 \leq 22.22$

$\hat D_2-s(\hat{D_2})*T \leq D_2 \leq \hat D_2+s(\hat{D_2})*T$, i.e. $-21.28 \leq D_2 \leq 24.08$

$\hat D_3-s(\hat{D_3})*T \leq D_3 \leq \hat D_3+s(\hat{D_3})*T$, i.e. $-20.81 \leq D_3 \leq 24.55$

We can see that the three CIs all includes zero, meaning that there are no significant difference between the means for three disk drive maker.

## (f)
$\hat \mu_{.1}=\bar Y_{.1.} = 56.13$ 

$c_1 = (3-1)/(5*3*3) = 2/45$, $c_2=1/(5*3*3)=1/45$

$s(\hat \mu_{.1}) = \sqrt{c_1*MSAB+c_2*MSA} = 3.712$

$df = s(\hat \mu_{.1})^4/((c_1*MSAB)^2/4+(c_2*MSA^2/2) = 4.16$

$t(0.995, 4)=4.6$

Therefore, using Satterthwaite procedure, the condidence interval is:

$\hat \mu_{.1}-t(0.995, 4)*s(\hat \mu_{.1}) \leq \mu_{.1} \leq \hat \mu_{.1}+t(0.995, 4)*s(\hat \mu_{.1})$, i.e. $39.05 \leq \mu_{.1} \leq 73.21$

It means that with confidence coefficient 0.99, we conclude that the the mean minutes for make 1 is somewhere between 39.05 and 73.21.


## (g)
$MS1=MSA$, $MS2=MSE$

$c1=1/15$,$c2=-1/15$

$\hat L=(MS1-MS2)/(5*3)=-2.648$

$df1=2$, $df2=36$

$F1=F(1-0.01/2,df1,\inf)=5.2983$

$F2=F(1-0.01/2,df2,\inf) = 1.7106$

$F3=F(1-0.01/2,\inf,df1)=199.4996$

$F4=F(1-0.01/2,\inf,df2)=2.0127$

$F5=F(1-0.01/2,df1,df2)=6.1606$

$F6=F(1-0.01/2,df2,df1)=199.4718$

$G1=1-1/F1=0.8113$

$G2=1-1/F2= 0.4154$

$G3=((F5-1)^2-(G1*F5)^2-(F4-1)^2)/F5 = 0.1022$ 

$G4=F6*((1-1/F6)^2-(F3/F6-1/F6)^2-G2^2) = -35.3895$

$H_L=\sqrt{(G1*c1*MS1)^2+((F4-1)*c2*MS2)^2-G3*c1*c2*MS1*MS2}=3.605$ 

$H_U=\sqrt{((F3-1)*c1*MS1)^2+(G2*c2*MS2)^2-G4*c1*c2*MS1*MS2}=162.73$

Therefore, using MLS procedure, the confidence interval is:

$\hat L - H_L \leq L \leq \hat L + H_U$

-2.648-3.605 $\leq$ L $\leq$ -2.648+162.73, i.e. -6.253 $\leq$ L $\leq$ 160.082

A negativa variance is not meaningful, so the CI can be written as [0, 160.082].

It means that with confidence coefficient 0.99, we conclude that the variability of mean minutes for different techinician is somewhere between 0 and 160.082.
 


