---
title: "STA141 Assignment 1 II"
author: "Zhen Zhang"
date: "October 4, 2015"
output: pdf_document
---

```{r global_options, include=FALSE, echo=FALSE, cache=FALSE}
options(width=60)
knitr::opts_chunk$set(warning=FALSE, message=FALSE, tidy=TRUE, fig.path='Figs/')
load("STA141/vehicles.rda")
library(ggplot2)
```

**Anomolies**

1. Look at the last observation:

```{r}
vposts[dim(vposts)[1],]
```

The price is only 5 dollars. That is strange. So look at the title, it says that 'asking 5k OBO', which means 5k dollars or best offer. So the price should be 5k.

For now, I will not stop. There are many observations with price 5 dollars

```{r}
priceIndex <- which(vposts$price == 5)
vposts$body[priceIndex]
```

From the result I can see these situations: best offer, trade, accessories and so on. So we need to pay attentation to the prices, looking at their description, and see what they are really talking about. Maybe a new variable can be created to reflect this kind of information.

2. Look at the distribution of the year:

```{r}
table(vposts$year)
```

There are some interesting values: 4 and 2022. It is obvious that they are not reasonable. For the value 4, I see the body reveals no message. When looking at the value 2022, I think '2022' maybe some part of the model. This reminds us that there maybe some other vehicles that are mistyped, for example, let's look at vehicles with year 2000-2009:

```{r}
vpostsYear <- vposts[vposts$year %in% seq(2000,2009),]
yearGsub <- as.numeric(gsub('.*(20[0-9]{2}).*','\\1',vpostsYear$body))
```

What I do here is first extract the dataframe with year in the range of 2000 and 2009, and then use regular expression to extract the year in the body. Now calculate how much difference between them:

```{r}
yearVector <- unlist(subset(vposts, year %in% seq(2000,2009), year))
names(yearVector) <- NULL
yearBadIndex <- which(yearVector != yearGsub)
length(yearBadIndex)
```

There are 2889 observations that are not equal. But this may due to the difference from the model year and the bought year, but we still need to pay attention to the differences. The correction can be done by substitute the value we get from gsub into the original vposts dataset.

3. First let's look at the plot between condition and odometer.

I need to extract some main conditions from the data: 

```{r}
condTable <- table(vposts$condition)[order(table(vposts$condition),decreasing = T)[1:8]]
vpostsCond <- vposts[vposts$condition %in% rownames(condTable),]
vpostsCond$condition <- factor(vpostsCond$condition)
```

```{r}
ggplot(vpostsCond,aes(x = condition,y = odometer)) + geom_boxplot() + scale_y_continuous(limits=c(0,50000))
```

It is easily seen that for vehicles with condition new, there are many values with odometer above 40000, even larger than some other conditions like used and fair. These may be a false judgement of the condition of the cars:

```{r}
newOdometerBadIndex <- which(vpostsCond$condition == 'new' & vpostsCond$odometer > 35000)
dim(subset(vpostsCond,condition == 'new' & odometer >35000))
```

The result suggests that there are 73 such values. I believe it may not be reasonable new vehicle to have odometers more than 40000. There may also be some circumstances that some other conditions are also falsely judged. This will have a heavy impact on the consumers, and for those data, the conditions need to be evaluated one more time.

But there is also a simple correction: for the observations that are identified as outliers of the condition new in the sense of odometer, just change them to used, since they fall outside the 25% to 75% range of all conditions except used from the plot.

**Interesting Features**

1. City plays an important role in vehicle manufacturing. The makers need to do some market research to confirm which model to produce, which price to set, which mpg to build on the basis on cities. Here I will dig into two features classified by cities: cylinders and prices.

+ *Cylinders*

Cylinders is an important feature. Let me first draw the plot for city and cylinders.

```{R}
ggplot(vposts,aes(x=city,y=cylinders)) + geom_boxplot()
```

The first finding is that, the cylinders for city boston, nyc, sac and sfbay are much the same, both have mean 6 and interval 4.5 to 6, and for city chicago, denver and lasvegas, they are the same as well. This means that, the saler may produce the same cars in type of cylinders in some specific cities.

+ *Price*

There is a phenomenon called price discrimination in economics, meaning that charging different fares to different groups of people to earn as much profit as possible. In the industry of vehicles, this principle also applies. Here is what I find:

Of course, we first need to remove the unreasonable values. There are four such values:

```{r}
tail(sort(vposts$price))
```

So I set the threshold as 9999998.

```{r}
vpostsPrice = vposts[vposts$price < 9999998, ]
by(vpostsPrice, vpostsPrice$city, function(x) {mean(x$price,na.rm = T)})
```

The output is obvious: The car price at denver and sfbay is larger than anywhere else. Taken into consideration that the numbers of vehicles are almost the same across the cities, It may because the prices there are high or the vehicles sold there is better than anywhere else. And the prices for sac, boston and nyc are the lowest. 

This tells us that, when selling cars, no matter you are a dealer or an owner, you should charge different prices for different cities to get your car sold in a reasonable price quickly.

I also use the plots to illustrate my point:

```{r}
ggplot(vposts, aes(x = price)) + geom_histogram() + scale_x_continuous(limit = c(0,3*10^4)) + facet_wrap(~city, scales = 'free')
```

2. I have done some analysis related to byOwner or byDealer in the previous homework. Now I want to do more analysis on this variable to reveal more information.

First let's see the year distribution:

```{r}
table(vposts$byOwner,vposts$year)
```

It is seen that, if you want to but a car before the year of 2004, then buy it from the owner. Otherwise, choose dealer.

```{r}
by(vpostsPrice, vpostsPrice$byOwner, function(x) {mean(x$price,na.rm = T)})
```

Here I also compare the price. As expected, the price of buying from owner is much lower than buying from dealer, almost only half the price.

So with lowe price, what about the condition? let me check:

Then categorize them:

```{r}
table(vpostsCond$condition,vpostsCond$byOwner)
```

Basically, the condition of Owners' are not as good as that of Dealers'. For example, if you want certified ones, you can only buy them from dealer.

```{r}
vpostsOdo <- vposts[vposts$odometer < 10^7,]
by(vpostsOdo, vpostsOdo$byOwner, function(x) {mean(x$odometer,na.rm = T)})
```

The last comes the odometer analysis. The odometer for owner is more than 30000 miles, about 40% of that of the dealer.

So I have given a more thorough analysis on the type by owner or by dealer.

3. There are some features that are not listed in this dataset. For example, the engine value, the mpg value and the color value. So let's first add them to the dataset.

How to add them on? The information can be read from the variable body. So first have a snapshot of the first observation:

```{r}
vposts$body[1]
```

When looking at the third line, I can see some features listed below:

+ VIN: 2G1FT1EW1C9106920

+ Stock #: 101083A

+ Condition: Used

+ Mileage: 16,324

+ MPG: 16

+ Exterior: Black

+ Interior: Black

+ Body: Coupe

+ Engine: 6.2L V8

So here come some important features: Condition(already in the dataset), MPG, Color(Exterior & Interior), Body(The vehicle type, already in the dataset) and Engine. I will create them and add them to the dataset now.

```{r}
## grep them from the dataset
MPG <- gsub('.*MPG: ([0-9]+).*','\\1',vposts$body)
Exterior <- gsub('.*Exterior: (.*)Interior.*','\\1',vposts$body)
Interior <- gsub('.*Interior: ([A-Z][a-z]+).*','\\1',vposts$body)
Engine <- gsub('.*([0-9][.][0-9])L.*','\\1',vposts$body)
## set the unchanged value to NA
MPG[MPG == vposts$body] <- NA
Exterior[Exterior == vposts$body] <- NA
Interior[Interior == vposts$body] <- NA
Engine[Engine == vposts$body] <- NA
## add them to the dataset
vposts$MPG <- as.numeric(MPG)
vposts$Exterior <- Exterior
vposts$Interior <- Interior
vposts$Engine <- as.numeric(Engine)
```

So now let's see the relationship between MPG and city:

```{r}
ggplot(vposts, aes(x = city, y = MPG)) + geom_boxplot(outlier.shape = NA) + scale_y_continuous(limits = c(10,40))
```

I control the limit because of some outliers. Then it is trivial from the plot that the city of denver has lowest MPG, while boston and nyc have the largest MPG.

What about the relationship between Engine and city?

```{r}
ggplot(vposts, aes(x = city, y = Engine)) + geom_boxplot(outlier.shape = NA)
```

Almost no clear pattern.