---
title: "STA135 homework6"
author: "Zhen Zhang"
date: "February 6, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy = TRUE)
options(width = 65)
```

## 5.10

```{r}
growth <- read.table("/Users/Elliot/Github/UCDavis/STA135/resource/STA135_data/T1-4.dat")
growth_length <- growth[,5:8]
names(growth_length) <- paste0("Length", 2:5)
```

### (a)

Simultaneous confidence intervals:

```{r}
calculate_simultaneous_interval <- function(sample_data, test_mean, alpha, p, flag = T, print_flag = T) {
  names_sample_data <- names(sample_data)
  sample_data <- as.matrix(sample_data)
  n <- dim(as.matrix(sample_data))[1]
  sample_mean <- apply(sample_data, 2, mean)
  deviation <- sample_mean - test_mean
  covariance_matrix <- cov(sample_data)
  F_star <- sqrt((n-1)*p/(n-p)*qf(1-alpha, p, n-p))
  sd_deviation <- sqrt(diag(covariance_matrix)/n)
  part2 <- F_star * sd_deviation
  result <- matrix(c(deviation - part2, upper = deviation + part2), n <- dim(as.matrix(sample_data))[2])
  rownames(result) <- names_sample_data
  colnames(result) <- c("lower", "upper")
  if (print_flag) print(result)
  if (flag) return(F_star)
  return(result)
}

invisible(calculate_simultaneous_interval(growth_length, 0, 0.05, 4))
```

### (b)

simultaneous intervals for changes in length

```{r}
growth_length_change <- as.data.frame(sapply(1:3, function(i) growth_length[,i+1] - growth_length[,i]))
names(growth_length_change) <- paste0("Length",3:5,"-Length",2:4)
invisible(calculate_simultaneous_interval(growth_length_change, 0, 0.05, 4))
```

### (c)

The simultaneous interval is:

```{r}
growth_length_compare <- as.data.frame(cbind(growth_length[, 2] - growth_length[, 1], growth_length[, 4] - growth_length[, 3]))
names(growth_length_compare) <- paste0("Length",c(3, 5),"-Length",c(2, 4))
invisible(calculate_simultaneous_interval(growth_length_compare, 0, 0.05, 4))
```

The ellipse:

```{r}
plot_ellipse <- function(sample_data, F_star, test_point) {
  n <- dim(sample_data)[1]
  sample_data_var <- cov(sample_data)
  sample_data_eigenvalue <- eigen(sample_data_var)$values
  main_vector <- eigen(sample_data_var)$vectors[,1]
  lambda1 <- sqrt(sample_data_eigenvalue)[1]
  lambda2 <- sqrt(sample_data_eigenvalue)[2]
  alpha <- atan(main_vector[2] / main_vector[1])
  theta <- seq(0, 2 * pi, length=(1000))
  F_star <- F_star / sqrt(n)
  a <- lambda1 * F_star
  b <- lambda2 * F_star
  x_bar <- unname(apply(sample_data, 2, mean))
  x <- x_bar[1] + a * cos(theta) * cos(alpha) - b * sin(theta) * sin(alpha)
  y <- x_bar[2] + a * cos(theta) * sin(alpha) + b * sin(theta) * cos(alpha)
  plot(x, y, type = "l")
  points(x_bar[1], x_bar[2], type = 'p')
  points(test_point[1], test_point[2], type = 'p', col = 'red')
}

plot_ellipse(growth_length_compare, calculate_simultaneous_interval(growth_length_compare, 0, 0.05, 4), c(-110,-110))
```

I will revisit the plot in part (e).

### (d)

```{r}
calculate_Bonferroni_interval <- function(sample_data, test_mean, alpha, pair, print_flag = T) {
  names_sample_data <- names(sample_data)
  sample_data <- as.matrix(sample_data)
  n <- dim(as.matrix(sample_data))[1]
  sample_mean <- apply(sample_data, 2, mean)
  deviation <- sample_mean - test_mean
  covariance_matrix <- cov(sample_data)
  B <- qt(1-alpha/(2*pair), n-1)
  sd_deviation <- sqrt(diag(covariance_matrix)/n)
  part2 <- B * sd_deviation
  result <- matrix(c(deviation - part2, upper = deviation + part2), n <- dim(as.matrix(sample_data))[2])
  rownames(result) <- names_sample_data
  colnames(result) <- c("lower", "upper")
  if (print_flag) print(result)
  return(result)
}

invisible(calculate_Bonferroni_interval(growth_length, 0, 0.05, 7))
invisible(calculate_Bonferroni_interval(growth_length_change, 0, 0.05, 7))
```


### (e)

The bonferroni interval:

```{r}
invisible(calculate_Bonferroni_interval(growth_length_compare, 0, 0.05, 7))

plot_ellipse(growth_length_compare, calculate_simultaneous_interval(growth_length_compare, 0, 0.05, 4, print_flag = F), c(-110,-110))
s_interval <- as.vector(t(calculate_simultaneous_interval(growth_length_compare, 0, 0.05, 4, FALSE, print_flag = F)))
b_interval <- as.vector(t(invisible(calculate_Bonferroni_interval(growth_length_compare, 0, 0.05, 7, print_flag = F))))
for (i in 1:2) {
  abline(v = s_interval[i], lty = 2)
  abline(v = b_interval[i], lty = 3)
}

for (i in 3:4) {
  abline(h = s_interval[i], lty = 2)
  abline(h = b_interval[i], lty = 3)
}
```

The bigger square with line type `lty = 2` is the simultaneous interval, while the inner square with line type `lty = 3` is the bonferroni interval. Clearly the bonferroni interval is much small.

# 5.20

```{r}
kites <- read.table("/Users/Elliot/Github/UCDavis/STA135/resource/STA135_data/T5-12.dat")
names(kites) <- c("tail length", "wing length")
```

### (a)

```{r}
plot_ellipse(kites, calculate_simultaneous_interval(kites, 0, 0.05, 2), c(190,275))
```

Here is the ellipse. The point is in the ellipse, so it is plausible.

### (b)

Simultaneous interval:

```{r}
print(calculate_simultaneous_interval(kites, 0, 0.05, 2))
```

Bonferroni interval:

```{r}
print(calculate_Bonferroni_interval(kites, 0, 0.05, 2))
```

The simultaneous interval is a little large than the bonferroni interval. The simultaneous confidence interval will touch the simultaneous confidence region from outside.

### (c)

```{r}
par(mfrow = c(1,3))
qqnorm(kites$`tail length`, main = "qqplot for X1")
qqnorm(kites$`wing length`, main = "qqplot for X2")
with(kites, plot(`tail length`, `wing length`))
```

From the plot we know $X_1$ and $X_2$ are not normal. We could transform $X_1$.