---
title: "STA206 ASS2"
author: "Zhen Zhang"
date: "October 9, 2015"
output: pdf_document
---

```{r global_options, include=FALSE, echo=FALSE, cache=FALSE}
options(width=60)
knitr::opts_chunk$set(warning=FALSE, message=FALSE, tidy=TRUE, fig.path='Figs/')
```

(a) First read the data into R, and then give the data the column names

```{r}
muscle <- read.table('~/Desktop/muscle.txt')
names(muscle) <- c('mass','age')
```

Next I use plot function to draw the scatterplot:

```{r}
with(muscle, plot(age, mass))
```

From the scatterplot, I can see a roughly tendancy that the amount of muscle mass decreases with age.

(b) First I need to load the MASS package which includes the box-cox function.

```{r}
library(MASS)
```

Then use the box-cox procedure

```{r}
boxcox(mass ~ age, data = muscle, lambda = seq(0,10,1))
```

From the plot, I see that a transformation is not needed: order 1, which is linear, is best to fit the data.

(c) Let me run the linear regression:

```{r}
lmfit <- lm(mass ~ age, data = muscle)
```

Now display the summary

```{r}
summary(lmfit)
```

The coefficients are $\hat{\beta_0}$ = 156.3466, $\hat{\beta_1}$ = -1.1900. The standard error for $\hat{\beta_0}$ is 5.5123, for $\hat{\beta_1}$ is 0.0902. MSE is 8.173, degree of freedom is 58.

(d) The regression line is:

$$mass_i = 156.3466 - 1.1900 * age_i$$

Now add it to the scatterplot:

```{r}
with(muscle, plot(age, mass))
abline(lmfit)
```

The line fits the data well.

(e) The fitted values and residuals for the 6th and 16th cases in the data set are:

```{r}
lmfit$fitted.values[c(6,16)]
lmfit$residuals[c(6,16)]
```

The fitted value and residual for the 6th case are 107.55675 and 11.443252 respectively.

The fitted value and residual for the 6th case are 90.89681 and -3.896811 respectively.

(f) Here we will see four plots, and the first one is the residuals vs. fitted values plot, the second one is the residuals Normal Q-Q plot.

```{r}
par(mfrow = c(2,2))
plot(lmfit)
```

The simple linear regression model with Normal errors:

$$mass_i = 156.3466 - 1.1900 * age_i + \epsilon_i$$

Assumptions:

+ The distribution of error terms, say $\epsilon_i$ are independent, identical normal distribution (i.i.d with Normal).

+ The relationship is linear.

Comment: The linear relationship holds, but the normal distribution is not valid. It is light tailed than normal distribution.

(g) first I need to know the t value for two-tailed 99% with degree of freedom 58:

```{r}
t1 <- qt(0.995, 58)
```

So the intercept is:

$$[\hat{\beta_0} - t(0.995, 58) * se(\hat{\beta_0}), \hat{\beta_0} + t(0.995, 58) * se(\hat{\beta_0})]$$

evaluated at:

$$[141.6657, 171.0274]$$

(h) Null hypothesis ($H_0$): There is no linear association between the amount of muscle mass and age.

Alternative hypothesis ($H_1$): There is a negative linear association between the amount of muscle mass and age.

The test statistic: T-statistic: $T^* = \frac{\hat{\beta_1}}{se(\hat{\beta_1})}$

The null distribution: Under $H_0$, $\beta_0$ = 0, $T^*$ ~ $t(58)$

The decision rule: Reject $H_0$ if and only if $T^*$ < $t(0.01,58)$

The conclusion: the t statistic here is $t(0.01,58)$ = -2.392377, and $T^*$ is -13.1929, so reject $H_0$, meaning there is a negative linear association between the amount of muscle mass and age.

(i) use the function predict in r:

```{r}
predict(lmfit, data.frame(age = 60), interval="predict")
```

So the prediction interval is $[68.45067, 101.443]$.

Interpretation: if I sample 60 women from the dataset, then the muscle mass with age 60 is 95% probability falls into the interval [68.45067, 101.443].

(j) Test for anova:

```{r}
anovafit <- anova(lmfit)
```

Null hypothesis ($H_0$): There is no linear association between the amount of muscle mass and age.

Alternative hypothesis ($H_1$): There is a linear association between the amount of muscle mass and age.

The test statistic: F-statistic: $F^* = \frac{MSR}{MSE}$

The null distribution: Under $H_0$, $F^*$ ~ $F(0.99,1,58)$

The decision rule: Reject $H_0$ if and only if $F^*$ > $F(0.99,1,58)$

The conclusion: the t statistic here is $F(0.99,1,58)$ = 7.093097, and $F^*$ is 174.06, so reject $H_0$, meaning there is a linear association between the amount of muscle mass and age.

(k) let's see 

```{r}
summary(lmfit)
```

Since the $r^2$ is 0.7501, then 75.01% of the total variation in muscle mass is “explained” by age.

The correlation coefficient between muscle mass and age is $-\sqrt{0.7501}$ = $-0.866$.